Project Uccle

### 2020-05-30 project.Rmd
cleanup project files

```{r "libraries", include=FALSE}
library(knitr)
opts_chunk$set(warning = F, message = F)
library(stringi) # %s+%
library(lubridate) # floor_date
library(xts) # zoo
library(fpp) # auto.arima
library(ggplot2)
# library(gridExtra) # for gg-plotting multiple plots
library(cowplot) # for plot_grid
# library(car) # qqPlot
library(qqplotr) # qqplot with ggplot
library(tidyverse) # map_df, read_csv, group_by, summarize
library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
```

@@

```{r cleanup, echo=FALSE}
out_domain <- (seis$min < domain[1] | seis$max > domain[2])
invalid_values <- (seis$min == -1 & seis$max == 1) | (seis$min == 0 & seis$max == 0) | out_domain
cat("Removing", sum(out_domain), "values out of domain,", (sum(invalid_values) - sum(out_domain)), "explicit NA\n")
seis <-
  seis %>%
  mutate(min=replace(min, invalid_values, NA), max=replace(max, invalid_values, NA)) %>%
  drop_na()
# %>% mutate(belgian_second=with_tz(utc_second, "Europe/Brussels"))
remove(out_domain, invalid_values) # free ~350 Mb from RAM
```

```{r missing values analysis}
seis.complete <- seis %>% complete(utc_second=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = as.difftime("0:0:1")))
seis.first.second <- min(seis$utc_second)
seis.missing.ids <- which(is.na(seis.complete$min))

@@

```{r missing values hour analysis}
seis.complete <- seis.h %>% complete(utc_hour=seq.POSIXt(from=min(seis.h$utc_hour), to=max(seis.h$utc_hour), by = as.difftime("1:0:0")))
seis.first.second <- min(seis$utc_second)
seis.missing.ids <- which(is.na(seis.complete$mean_movement_s))

@@

@@

```{r periodgram}
library(spec)
pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end), main="Periodgram", detrend=F, demean=T, na.action=na.aggregate, plot=F, xlim=c(0,1/3600))
pgram.data <- data.frame(freq=pgram$freq, spec=pgram$spec)
ggplot(pgram.data) +
  geom_line(aes(freq, log(spec))) +
  xlab("frequency") + ylab("log(spectrum)") +
  labs(title = "Periodgram", caption = paste0("bandwidth = ", pgram$bandwidth, "\tdf = ", pgram$df)) ->
  periodgram

@@

ratio.color <- function(x) ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))

library(ggrepel)
periodgram +
  scale_x_continuous(sec.axis = sec_axis(~1/./3600, breaks=c(1, 6*c(1,2,4,4*7)), labels = as.integer, name="hours")) +
  scale_y_continuous(sec.axis = sec_axis(~exp(.), breaks=mean(pgram$spec) + sd(pgram$spec) * c(1:3), labels=c("1", "2", "3")[1:3], name="spectrum/sd ratio")) +
  geom_hline(yintercept=log(mean(pgram$spec) + sd(pgram$spec) * c(1:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(1:3))) +
  geom_point(data=data.frame(freq=top95freq, spec=top95pc), aes(freq, log(spec), color=ratio.color(spec)), shape="o", size=4, show.legend = F) +
  geom_label_repel(data=data.frame(freq=top95freq, spec=top95pc)[1:3,], aes(freq, log(spec), label=round(1/freq/3600), color=ratio.color(spec)), segment.color="grey", show.legend=F) +
  scale_color_manual(values = c("darkgreen", "blue", "red")[start_ratio:3])

@@

@@

```
Once notice the composition is additive (the fluctuations does not seem to grow with the trend)
```{r decompose}
dec.addi <- decompose(seasonality.aggregate(seis.h.train))
autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period))
dec.multi <- decompose(log(seasonality.aggregate(seis.h.train)))
autoplot(dec.multi) + labs(title="Decomposition of multiplicative time series (log-scale additive)", subtitle = paste0("Seasonality ", seasonality.period))
dec.multi.plain <- decompose(seasonality.aggregate(seis.h.train), type="multip")
autoplot(dec.multi.plain) + labs(title="Decomposition of multiplicative time series", subtitle = paste0("Seasonality ", seasonality.period))
plot_grid(
  autoplot(acf(dec.addi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  autoplot(acf(dec.multi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  autoplot(acf(dec.multi.plain$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  nrow = 3)
plot_grid(
  autoplot(pacf(dec.addi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  autoplot(pacf(dec.multi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  autoplot(pacf(dec.multi.plain$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  nrow = 3)
```

We now extract the trend using a linear filter with smoothing parameter $p$ equal to 180 (almost half a year):

@@

@@

>>>>

```{r "libraries", include=FALSE}
library(knitr)
opts_chunk$set(warning = F, message = F)
library(stringi) # %s+%
library(lubridate) # floor_date
library(xts) # zoo
library(fpp) # auto.arima
library(ggplot2)
library(ggrepel) # for nicer labels on ggplot
# library(gridExtra) # for gg-plotting multiple plots
library(cowplot) # for plot_grid
# library(car) # qqPlot
library(qqplotr) # qqplot with ggplot
library(spec) # for periodgram
library(tidyverse) # map_df, read_csv, group_by, summarize...
library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
```

@@

```{r cleanup, echo=FALSE}
out_domain <- (seis$min < domain[1] | seis$max > domain[2])
invalid_values <- (seis$min == -1 & seis$max == 1) | (seis$min == 0 & seis$max == 0) | out_domain
cat("Removing", sum(out_domain), "values out of domain,", (sum(invalid_values) - sum(out_domain)), "explicit NA\n")
seis <-
  seis %>%
  mutate(min=replace(min, invalid_values, NA), max=replace(max, invalid_values, NA)) %>%
  drop_na()
# %>% mutate(belgian_second=with_tz(utc_second, "Europe/Brussels"))
remove(out_domain, invalid_values) # free ~350 Mb from RAM
```
## Missing values analysis (seconds)
```{r missing values analysis}
seis.complete <- seis %>% complete(utc_second=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = as.difftime("0:0:1")))
seis.first.second <- min(seis$utc_second)
seis.missing.ids <- which(is.na(seis.complete$min))

@@

## Missing values analysis (hours)
```{r missing values hour analysis}
seis.complete <- seis.h %>% complete(utc_hour=seq.POSIXt(from=min(seis.h$utc_hour), to=max(seis.h$utc_hour), by = as.difftime("1:0:0")))
seis.first.second <- min(seis$utc_second)
seis.missing.ids <- which(is.na(seis.complete$mean_movement_s))

@@

# Analysis
```{r whole data plot}
month.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'month'), to=ceiling_date(seis.h.end, 'month'), by = "1 month")
hours.from.start <- as.numeric(month.starts-seis.h.start)*24+168
autoplot(seis.h.ts, type="l") +
  scale_x_continuous(breaks=hours.from.start/168, labels=format(month.starts, "%m-%y")) +
  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in original scale")
dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period)) + xlab("Weeks")
```

@@

## Periodgram
```{r periodgram}
pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end), main="Periodgram", detrend=F, demean=T, na.action=na.aggregate, plot=F, xlim=c(0,1/3600))
pgram.data <- data.frame(freq=pgram$freq, spec=pgram$spec)
ggplot(pgram.data) +
  geom_line(aes(freq, log(spec))) +
  xlab("frequency") + ylab("log(spectrum)") +
  labs(title = "Periodgram", caption = paste0("bandwidth = ", pgram$bandwidth, "\tdf = ", pgram$df)) ->
  periodgram

@@

ratio.color <- function(x) ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
shown_labels <- 4
periodgram +
  scale_x_continuous(sec.axis = sec_axis(~1/./3600, breaks=c(1, 6*c(1,2,4,4*7)), labels = as.integer, name="hours")) +
  scale_y_continuous(sec.axis = sec_axis(~exp(.), breaks=mean(pgram$spec) + sd(pgram$spec) * c(1:3), labels=c("1", "2", "3")[1:3], name="spectrum/sd ratio")) +
  geom_hline(yintercept=log(mean(pgram$spec) + sd(pgram$spec) * c(1:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(1:3))) +
  geom_point(data=data.frame(freq=top95freq, spec=top95pc), aes(freq, log(spec), color=ratio.color(spec)), shape="o", size=4, show.legend = F) +
  geom_label_repel(data=data.frame(freq=top95freq, spec=top95pc)[1:shown_labels,], aes(freq, log(spec), label=round(1/freq/3600), color=ratio.color(spec)), segment.color="grey", show.legend=F) +
  scale_color_manual(values = c("darkgreen", "blue", "red")[start_ratio:3])

@@

## Decomposition

@@

```
Once notice the composition is additive (the fluctuations does not seem to grow with the trend)

```{r decompose}
dec.addi <- decompose(seasonality.aggregate(seis.h.train))
autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period))
dec.multi <- decompose(log(seasonality.aggregate(seis.h.train)))
autoplot(dec.multi) + labs(title="Decomposition of multiplicative time series (log-scale additive)", subtitle = paste0("Seasonality ", seasonality.period))
dec.multi.plain <- decompose(seasonality.aggregate(seis.h.train), type="multip")
autoplot(dec.multi.plain) + labs(title="Decomposition of multiplicative time series", subtitle = paste0("Seasonality ", seasonality.period))
plot_grid(
  autoplot(acf(dec.addi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  autoplot(acf(dec.multi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  autoplot(acf(dec.multi.plain$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  nrow = 3)
plot_grid(
  autoplot(pacf(dec.addi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  autoplot(pacf(dec.multi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  autoplot(pacf(dec.multi.plain$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
  nrow = 3)
```

### Manual filtering/decomposition
We now extract the trend using a linear filter with smoothing parameter $p$ equal to 180 (almost half a year):

@@

### Multi trend extraction

@@

# Forecast

### 2020-05-30 project.mems.Rmd
cleanup project files

```{r tentatives, eval=FALSE, child='chapter1.Rmd', include=FALSE}
```

@@

```{r arima}
to_test <- list(
  c(0,1,0,0,1,1)
)

@@

```{r arima on seasonality}
autoplot(residuals+trend.s)
residuals.bc.lambda <- BoxCox.lambda(residuals+trend.s)
residuals.bc <- BoxCox(residuals+trend.s, residuals.bc.lambda)
residuals.bc %>% ggtsdisplay(lag.max = 168*5)
residuals.bc %>% diff() %>% ggtsdisplay(lag.max = 168*5)
residuals.bc %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
residuals.bc %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
residuals.bc %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
residuals.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
residuals.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168) # arima (0,1,0)(0,1,1)[168]
# acf cuts off after p=168, pacf shows peaks each 168
residuals.bc %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)

@@

>>>>

@@

## Arima
```{r arima}
to_test <- list(
  c(0,1,0,0,1,1)
)

@@

## Arima without seasonality
```{r arima on seasonality}
autoplot(residuals+trend.s)
residuals.bc.lambda <- BoxCox.lambda(residuals+trend.s)
residuals.bc <- BoxCox(residuals+trend.s, residuals.bc.lambda)
residuals.bc %>% ggtsdisplay(lag.max = 168*5)
residuals.bc %>% diff() %>% ggtsdisplay(lag.max = 168*5)
residuals.bc %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
residuals.bc %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
residuals.bc %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
residuals.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
residuals.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168) # arima (0,1,0)(0,1,1)[168]
# acf cuts off after p=168, pacf shows peaks each 168
residuals.bc %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)

@@

## MSTL

### 2020-05-30 project.uccs.Rmd
working on draft
cleanup project files

@@

```{r mstl on msts}
seis.h.train.m <- msts(seis.h.train, c(24, 168))
seis.mstl <- seis.h.train.m %>% mstl(s.window="periodic") #c(15, 7))
seis.mstl.fc <- forecast(seis.mstl, h=168*18)

@@

```{r eval=FALSE, include=FALSE}
gc()
if (name == "MEM") {
  proj_file <- 'proj.mems.RData'
  save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
  remove(proj_file)
}
```

>>>>

## Arima

@@

## MSTL
```{r mstl on msts}
seis.h.train.m <- msts(seis.h.train, c(24, 168))
seis.mstl <- seis.h.train.m %>% mstl(s.window="periodic") #c(15, 7))
seis.mstl.fc <- forecast(seis.mstl, h=168*18)

@@

```{r save RData, include=FALSE}
gc()
save.image(proj_file)
```
